An Empirical Study of Work Fragmentation in Software Evolution Tasks

### 2015-08-07 paper.tex
discussion, threats, conclusions

\section{Comparison of results with Mylyn and UDC}
Once the UDC data was similar to the format of the Mylyn data we executed the same methods to have an answer to the proposed research questions with this new dataset.

@@

\section{RQ5: Common events during Recovery Time}
To better understand the reasons of the difference between positive and negative interruptions, we did an analysis of the common events during Recovery Time, depending on the nature of the interruption. For this we created a new classification of events on a lower level that takes into consideration what kind of activity is performed by the programmer. The new classification contains the categories described on the Table \ref{tbl:new_events_udc}.

@@

It is noticeable the relation between type of events. For example is natural to think that if there is a lot of debugging events there is also a lot of clean and build events as they are necessary to begin a debugging session. There is also a link between high level navigation and debugging because navigating into a method may take you to another class.

@@

@@

>>>>

\section{Comparison of results with UDC}
Once the UDC data was similar to the format of the Mylyn data we executed the same methods to have an answer to the proposed research questions with this new dataset.

@@

\section{RQ5: Common events during Recovery Time}

To better understand the reasons of the difference between positive and negative interruptions, we did an analysis of the common events during Recovery Time, depending on the nature of the interruption. For this we created a new classification of events on a lower level that takes into consideration what kind of activity is performed by the programmer. The new classification contains the categories described on the Table \ref{tbl:new_events_udc}.

@@

It is noticeable the relation between type of events. For example is natural to think that if there is a lot of debugging events there is also a lot of clean and build events as they are necessary to begin a debugging session. There is also a link between high level navigation and debugging because navigating into a method may take you to another class.
\section{Discussion}
Although they represent the same kind of information both Mylyn and UDC datasets have some peculiarities that make them unique and require different treatments.

On one side we have Mylyn a low level monitor to capture interaction data to better fix bugs for Eclipse projects. And in the other hand Usage Data Collector, a low level events monitor as well but with a more general approach. As the latter was installed by default on all the Eclipse distributions from 2008 to 2010, the size of the information collected is several times bigger than Mylyn. This makes of the UDC data a better source of information to analyze phenomenons on a wide scale, in comparison with Mylyn. For example, with UDC we can have interaction data from Java, Web, and mobile developers, and more specific roles like back-end developers, testers and C++ users. We can also have more details about every event which allows us to imply more things well beyond the role or language of preference of the programmer.

Begin installed by default is also a disadvantage for UDC as many of the data is actually garbage. For example, there are a lot of users that have only a few events registered, or the work sessions tell us very little about the actual coding tendencies of the programmer. With Mylyn we can be sure that most of the data was captured under the assertiveness of having interaction data logged for further analysis.

On this particular scenario, an advantage of Mylyn is that we had a predefined classification of the events and saved this process made with UDC. But if we expect to do a low level classification like the one on the analysis of common events during Recovery Time, then UDC is a better option. This is probably one of the causes in the difference of effects on the results (more pronounced with Mylyn) because we used a different criteria to chose between Edition and Selection when classifying the events.

@@

The same threats about the metrics, implications and statistical conclusions apply on the results with UDC. Additionally, we define as a thread the classification of events for it was based on our own criteria of what should be an Edition and Selection event. We took this liberty due to the extra information we had, in comparison with Mylyn, and the actual lack of a pre classification. We had special care when choosing the proper classification but it depends on the definition on what is Edition and Selection for it is a factor that can alter the final results. However, we had a very clear boundary between both types and were consistent on our judgments.

Similar to the above, we faced a problem when dealing with the workbench activation and deactivation events for they can me executed under many circumstances. We performed live test to understand how this events work and after that we had some assurances about their operation, for example they mark when the user leaves and returns to Eclipse, but we had more uncertainty on other scenarios. Therefore, we decided to exclude them from the results and only us them to define an interruption or the end of a session.

Another thread is the random selection of users to get their data. Although a posterior filter allows us to manage with only valid sessions, we understand that this kind of data selection might be irresponsible. For future work we'll have another approach and be more selective on the information we work.

@@

With the Usage Data Collector dataset, we performed the same methods to compare the results obtained with Mylyn. We found the same observations on the effect of interruptions on productivity, and even though the effect is less marked we can reach the same conclusions.

We also analyzed the Recovery Time on detail, finding a higher productivity when the interruption is positive and we identified a positive evolution of the metrics as the programmer advances through this stage, reaching a peak of productivity between the last minutes of the Recovery Time and 10 minutes before the next interruption happens.

Moreover, we found that after a positive interruption the most common events during the first minutes of the Recovery Time are those with an assertive intention like edition, refactoring and saving a file. In contrast, after a negative interruption the most common events are debugging, searching and navigation around classes, which tell us about a doubtful and uncertain behavior from the programmer.
